Wigglyhedra

### 2025-01-04 MZ/answerReferee.tex

{\bf Required Corrections}

\begin{itemize}
\item \textsl{\color{gray} Remark 9: I believe there are (edge) cases of pseudotriangles with only 3 points on its boundary; e.g.~take pseudotriangles in the $n = 1$ case (unless I’m supposed to count one of the boundary point twice, which isn’t clear to me).} \\
We indeed count points with multiplicity. We defined corners and hinges by the sentence ``Two consecutive wiggly arcs along the boundary of~$c$ define a \emph{corner} (resp.~a \emph{hinge}) of~$c$ if they bound a convex (resp.~concave) angle of~$c$'' in Definition 6. So two corners can indeed coincide at the same point. To insist on this fact, we have added ``Note that two corners of~$c$ can coincide at the same point~$i \in [n]$'' in Definition 6 and ``(two corners can coincide at the same point~$i \in [n]$)'' twice in Remark 9. Thank you for pointing out that this was unclear.

\item \textsl{\color{gray} Definition 34: I believe the other label should be $2h$ instead of $2h+1$.} \\
You are absolutely right, this is a typo. Thank you for noticing it.

\item \textsl{\color{gray} $g$-vectors, §3.2: This is probably the most important one. I was very confused after reading your definition of a $g$-vector; in particular, whether a $g$-vector is actually a vector in $\mathbb{H}_{2n}$. In Definition 47, you define the $g$-vector as the orthogonal projection of $\hat{g}$ onto $\mathbb{H}_{2n}$. However, the sentence right after the defining equations for $\alpha^\pm$ seems to agree with the definition that one gets before taking the projection (i.e. supposedly $\hat{g}$). I accepted this as a typo (that the $g$ should’ve been a $\hat{g}$ here), but then looking through the $g$-matrices examples that you’ve provided in the figures, I found that most of their $g$-vectors (columns) do not live in $\mathbb{H}_{2n}$ and behave more like $\hat{g}$ instead – in fact, the entries of the $g$-vectors generally did not sum to $0$, which I believe is the definition of being a vector in $\mathbb{H}_{2n}$. In reading the rest of the paper I deduced that you do want $g$-vectors to be in $\mathbb{H}_{2n}$, but please make sure that it is consistent with the examples you’ve provided.} \\
We agree with the referee that the implicit identification of the $\hat{g}$-vectors and $g$-vectors in our illustrations was really confusing. We have proceed to the following changes:
\begin{itemize}
\item We have split Figure 12 into two figures. Figure 13 now presents the $\hat{g}$-vectors and the \mbox{$g$-vectors} of a wiggly pseudotriangulation side to side, so that the projection is clearer. Figure~14 presents the $g$-vectors and the $c$-vectors of a wiggly pseudotriangulation side to side, so that the duality is visible.
\item We made clear in Example 48 and the caption of Table 1 that we represent here the $\hat{g}$-matrices. We decided to keep the $\hat{g}$-matrices rather than the $g$-matrices as the former are easier to read from Figure 5, while the latter have unnecessarily involved coefficients due to the projection. We believe that representing the $g$-matrices here would not help the reader.
\end{itemize}

@@

\item \textsl{\color{gray} pg.\,6 (C2): Since $D$ is by definition a set of wiggly arcs, $c \in D$ can be confused with arcs in $D$ rather than connected components.} \\
Again, this was too implicit. The $c$ in all sums of this paragraph actually ranges over the wiggly cells of~$D$. We have added ``denoting by~$C$ the set of wiggly cells of~$D$'' and replaced~``$c \in D$'' by~``$c \in C$'' as indices of the sums.

\item \textsl{\color{gray} pg.\,6 (C3): Aren’t there examples of internal arcs which only bounds one wiggly cell? Take again $n = 1$ and the unique internal wiggly arc of any pseudotriangulation.} \\
We count with multiplicities. We have added ``(we count a wiggly arc~$\alpha$ with multiplicity~$2$ if it appears twice along the boundary of~$c$)'' to make it clear.

\item \textsl{\color{gray} pg.\,7 (C1): This is a personal tangential comment: is this a new sequence that is not currently in OEIS? If so it might be worth submitting it.} \\
Indeed, we should include it. It is not in OEIS at the moment.

\item \textsl{\color{gray} pg.\,7 (C2): I was a bit slow in understanding this; I actually had to go to the corresponding description for wiggly permutations and work my way backwards to understand this. I believe an example would help a lot.} \\
This is actually the way this decomposition was found: we first worked it out on wiggly permutations and transported it back to wiggly pseudotriangulations through the bijection.
We agree that examples will help here. We have decided to include the following:
\begin{itemize}
\item In Remark 14, we have added Figure 6 with some examples of wiggly pseudotriangulations obtained by added one more wiggly pseudotriangle.
\item In Remark 28, we have added an example listing all wiggly permutations of $[10]$ obtained from the wiggly permutation~$78621354$ of~$[8]$.
\item We also extended Remark 41 to explicitly point to these particular examples.
\end{itemize}

@@

\item \textsl{\color{gray} pg.\,8 (C1): Perhaps more precisely, “compatible with wiggly arcs in c”.} \\
Indeed, we have added ``compatible with all wiggly arcs of~$c$''.

\item \textsl{\color{gray} pg.\,9 (C1): Maybe this is standard abuse of notation, but technically you’ve provided the clique graph and not the complex itself.} \\
You are right, except that we provide here a little more than the graph as we embed the full complex in the plane. To make it explicit, we have added the sentence ``We have embedded the wiggly complex~$\mathrm{WC}_2$ in the plane, so that all triangles (including the external one) form a face of~$\mathrm{WC}_2$.''.

@@

\item \textsl{\color{gray} pg.\,12 (C1): This should be ``admissible gaps''?} \\
Thanks! Fixed.

\item \textsl{\color{gray} pg.\,13 (C1): I find this description confusing. Maybe ``Following in a counterclockwise direction...'' would be better.} \\
We have replaced by ``Following each wiggly pseudotriangle of~$T$ in counterclockwise direction''.
\item \textsl{\color{gray} pg.\,13 (C2) Definition 34*: $2h + 1$ should be $2h$.} \\
You are absolutely right, we have corrected this typo.

\item \textsl{\color{gray} pg.\,14 (C1): I think it is important to first point out that every $j \in [n]$ is the hinge of exactly one pseudotriangle, which implies the highlighted sentence.} \\
Indeed, we have added ``As every~$j \in [n]$ is the hinge of precisely one pseudotriangle of~$T$''.

\item \textsl{\color{gray} Figures in §3.2 and §3.3*: As mentioned in the previous section, all $g(T)$ in figures should be replaced by the correct $g$-matrices (whose column vectors are in~$\mathbb{H}_{2n}$).} \\
We now have the $\hat{g}$-matrix and $g$-matrix side by side in Figure 13. For Table~1, we have chosen to keep the $\hat{g}$-matrices (but we made it clear in the caption). See answer above for more explanations.

### 2025-01-04 draft/wigglyhedra.tex
more corrections

\begin{remark}
\label{rem:numberWigglyPseudotriangulations}
The number~$wp_n$ of wiggly pseudotriangulations of~$n+2$ points is given by
\[
\begin{array}{c|ccccccccc}
n & 1 & 2 & 3 & 4 & 5 & 6 & 7 & 8 & \dots \\
\hline
wp_n & 2 & 14 & 176 & 3232 & 78384 & 2366248 & 85534176 & 3602770400 & \dots
\end{array}
\]
To compute~$wp_n$, denote by $wp_n(x)$ the polynomial where the coefficient of~$x^i$ is the number of wiggly pseudotriangulations of~$n+2$ points with~$i$ \defn{final} internal wiggly arcs (\ie ending at the last point~$n+1$).
For instance,
\begin{align*}
%	wp_0(x) & = 1, \\
	wp_1(x) & = x + 1, \\
	wp_2(x) & = 3 x^3 + 5 x^2 + 4 x + 2, \\
	wp_3(x) & = 15 x^5 + 35 x^4 + 44 x^3 + 40 x^2 + 28 x + 14, \\
	wp_4(x) & = 105 x^7 + 315 x^6 + 520 x^5 + 630 x^4 + 620 x^3 + 514 x^2 + 352 x + 176.
\end{align*}
We invite the reader to check the expressions of~$wp_1(x)$ and~$wp_2(x)$ with \cref{exm:allSmallWigglyPseudotriangulations}.
We use this additional variable~$x$ to obtain a recursive formula for~$wp_n(x)$.
%Indeed, any wiggly pseudotriangulation on~$n+1$ points can be obtained from a wiggly pseudotriangulation on~$n+2$ points by collapsing the points~$n$ and~$n+1$.
Indeed, any wiggly pseudotriangulation of~$n+2$ points can be obtained from a wiggly pseudotriangulation of~$n+1$ points by inserting a single wiggly pseudotriangle with a hinge at~$n$.
%There are two kinds of wiggly pseudotriangulations of~$n+2$ points:
%\begin{itemize}
%\item those with the wiggly arc~$(n, n+1, \varnothing, \varnothing)$ are obtained from a wiggly pseudotriangulation of~$n+1$ points by choosing the position of the only corner of~$t_n$ which is not at~$n+1$,
%\item those without the wiggly arc~$(n, n+1, \varnothing, \varnothing)$ are obtained from a wiggly pseudotriangulation of~$n+1$ points by choosing the positions of the two corners of~$t_n$ which are not at~$n+1$.
%\end{itemize}
Each wiggly pseudotriangulation of~$n+1$ points with $i$ final internal wiggly arcs contributes to:
\begin{itemize}
\item $i+2$ wiggly pseudotriangulations of~$n+2$ points with $i+2$ final internal wiggly arcs, containing the wiggly arc~$(n, n+1, \varnothing, \varnothing)$,
\item $j+1$ wiggly pseudotriangulations of~$n+2$ points with $j$ final internal wiggly arcs, not containing the wiggly arc~$(n, n+1, \varnothing, \varnothing)$,  for all~$0 \le j \le i+1$.
\end{itemize}
\pagebreak
This translates to the recursive formula %for all~$n \ge 1$,
\begin{align*}
wp_{n}(x) & = \overbracket[.5pt]{x \cdot \frac{\partial}{\partial x} \big( x^2 wp_{n-1}(x) \big)}^{\substack{\text{wiggly pseudotriangulations} \\ \text{containing } (n, n+1, \varnothing, \varnothing)}} + \overbracket[.5pt]{\frac{\partial}{\partial x}\Big( \smash{\frac{wp_{n-1}(1)-x^3wp_{n-1}(x)}{1-x}} \Big)}^{\substack{\qquad\; \text{wiggly pseudotriangulations} \qquad\;  \\ \text{not containing } (n, n+1, \varnothing, \varnothing)}} \\
& = \frac{1}{(1 - x)^2} \big( wp_{n-1}(1) + x^2 (2 x^2 - 2 x - 1) wp_{n-1}(x) + x^4 (x - 1) wp_{n-1}’(x) \big).
\end{align*}
This enables us to quickly compute~$wp_n(x)$ and we obtain the numbers of wiggly pseudotriangulations~$wp_n$ by evaluating~$wp_n(x)$ at~$x = 1$.
%This enables us to quickly compute the generating series of the number of wiggly pseudotriangulations
%\[
%\sum_{n \ge 0} wp_n(1) y^n = 1 + 2 y + 14 y^2 + 176 y^3 + 3232 y^4 + 78384 y^5 + 2366248 y^6 + 85534176 y^7 + 3602770400 y^8 + 
%% 173300710720 y^9 + 9373542317760 y^10 + 563142033172480 y^11 + 37206559614499840 y^12 + 2681213937595142656 y^13 +
%\dots
%\]
\end{remark}

@@

We are now ready to consider the wiggly flip graph and the wiggly complex.
The following definition is illustrated in \cref{fig:wigglyComplex}\,(right) when~$n = 2$.
%
\begin{figure}
\centerline{\includegraphics[scale=1.1]{wigglyComplex}}
\caption{The wiggly complex~$\wigglyComplex_2$ (left) and the wiggly flip graph~$\wigglyFlipGraph_2$ (right).}
\label{fig:wigglyComplex}
\end{figure}

@@

\begin{remark}
\label{rem:numberWigglyPermutations}
The number~$wp_n$ of wiggly permutations of~$[2n]$ is given by
\[
\begin{array}{c|ccccccccc}
n & 1 & 2 & 3 & 4 & 5 & 6 & 7 & 8 & \dots \\
\hline
wp_n & 2 & 14 & 176 & 3232 & 78384 & 2366248 & 85534176 & 3602770400 & \dots
\end{array}
\]
To compute~$wp_n$, denote by $wp_n(x)$ the polynomial where the coefficient of~$x^i$ is the number of wiggly permutations of~$[2n]$ with $i$ \defn{admissible gaps} (\ie gaps~$\gamma$ between two consecutive positions such that there is no~$j \in [n]$ such that the value~$2j$ appears before the gap~$\gamma$ while the value~$2j-1$ appears after the gap~$\gamma$).
For instance,
\begin{align*}
%	wp_0(x) & = 1, \\
	wp_1(x) & = x + 1, \\
	wp_2(x) & = 3 x^3 + 5 x^2 + 4 x + 2, \\
	wp_3(x) & = 15 x^5 + 35 x^4 + 44 x^3 + 40 x^2 + 28 x + 14, \\
	wp_4(x) & = 105 x^7 + 315 x^6 + 520 x^5 + 630 x^4 + 620 x^3 + 514 x^2 + 352 x + 176.
\end{align*}
We invite the reader to check the expressions of~$wp_1(x)$ and~$wp_2(x)$ with \cref{exm:allSmallWigglyPermutations}.
Any wiggly permutation of~$[2n]$ can be obtained from a wiggly permutation of~$[2n-2]$ by inserting the last values~$2n-1$ and~$2n$.
Each wiggly permutation of~$[2n-2]$ with $i$ admissible gaps contributes to:
\begin{itemize}
\item $i+2$ wiggly permutations of~$[2n]$ with $i+2$ admissible gaps and where~$2n-1$ appears before~$2n$ (they must appear consecutively),
\item $j+1$ wiggly permutations of~$[2n]$ with $j$ admissible gaps and where $2n$ appears before~$2n-1$, for all~$0 \le j \le i+1$.
\end{itemize}
We thus obtain the same recursive equation on~$wp_n(x)$ as in \cref{rem:numberWigglyPseudotriangulations}.
\end{remark}

@@

Finally, consider a cover relation~$\sigma \lessdot \sigma^j$ in the wiggly lattice~$\wigglyLattice_n$ as described in~\cref{lem:wigglyCoverRelation}.
Using the notations of~\cref{lem:wigglyCoverRelation}, observe that
\begin{itemize}
\item $\alpha(\sigma^j, \ell) = \alpha(\sigma, \ell)$ for all~$\ell < i$ or~$\ell > k$,
\item $\alpha(\sigma^j, \ell) = \alpha(\sigma, \ell+k-j)$ for all~$i \le \ell < j$, and
\item $\alpha(\sigma^j, \ell) = \alpha(\sigma, \ell-j+i-1)$ for all~$j < \ell \le k$.
\end{itemize}
Hence, $\Psi(\sigma) \ssm \{\alpha(\sigma, j)\} = \Psi(\sigma^j) \ssm \{\alpha(\sigma^j, j)\}$.
This implies that~$\Psi$ is a directed graph morphism from the Hasse diagram of the wiggly lattice~$\wigglyLattice_n$ to the wiggly increasing flip graph~$\wigglyIncreasingFlipGraph_n$.
As the underlying undirected graphs are both regular of degree~$2n-1$, we conclude that~$\Phi$ and~$\Psi$ induce directed graph isomorphisms between the wiggly increasing flip graph~$\wigglyIncreasingFlipGraph_n$ and the Hasse diagram of the wiggly lattice~$\wigglyLattice_n$.
\end{proof}

@@

\begin{remark}
The map~$\Phi$ sends wiggly pseudotriangulations with $i$ final internal wiggly arcs to wiggly permutations with $i$ admissible gaps, explaining the similarities of \cref{rem:numberWigglyPseudotriangulations,rem:numberWigglyPermutations}.
\end{remark}

@@

\section{Wiggly fan and wigglyhedron}
\label{sec:geometry}

@@

Consider now a complete simplicial fan~$\Fan$ of~$\R^d$ with rays~$(\b{r}_v)_{v \in V}$ and cones~$\R_{\ge 0} \b{R}_\triangle$ for~${\triangle \in \Delta}$, where~$\b{R}_\triangle \eqdef \set{\b{r}_v}{v \in \triangle}$ as in \cref{prop:characterizationFan}.
To realize the fan~$\Fan$, one can try to pick a height vector~$\b{h} \eqdef (h_v)_{v \in V} \in \R^V$ and consider the polytope
\(
\polytope{P}_{\b{h}} \eqdef \set{\b{x} \in \R^d}{\dotprod{\b{r}_v}{\b{x}} \le h_v \text{ for all } v \in V}.
\)
The following classical statement characterizes the height vectors~$\b{h}$ for which the fan~$\Fan$ is the normal fan of this polytope~$\polytope{P}_{\b{h}}$.
We borrow the formulation from~\cite[Lem.~2.1]{ChapotonFominZelevinsky}.

@@

We denote by~$(\b{e}_i)_{i \in [d]}$ the standard basis of~$\R^d$.
For~$I \subseteq [d]$, we define~$\one_I \eqdef \sum_{i \in I} \b{e}_i$, and we often shorten~$\one_{[d]}$ by~$\one_d$.
We denote by~$\HH_d$ the hyperplane of~$\R^d$ defined by the equation~$\dotprod{\b{x}}{\one_d} = 0$, and we denote by~$\pi : \R^d \to \HH_d$ the orthogonal projection on~$\HH_d$, that is~$\pi(\b{x}) \eqdef \b{x} - (\dotprod{\b{x}}{\one_d} / d) \one_d$.

@@

\begin{figure}
	\centering
	\begin{tikzcd}
		1 \arrow[bend left]{r} & 2 \arrow[bend left]{r}\arrow[bend left]{l}& \cdots \arrow[bend left]{r}\arrow[bend left]{l}& m \arrow[bend left]{l}
	\end{tikzcd}
	\caption{The quiver \(Q_m\). The algebra \(A_m\) is a quotient of the path algebra \(kQ_m\).}
	\label{fig:am-quiver}
\end{figure}

@@

This work started during the \emph{Winterbraids XIII: School on low dimensional topology, braid groups and interactions}. We are grateful to the organizers Bérénice Delcroix-Oger, Peter Feller, Vincent Florens, Hoel Queffelec for this fruitful event.
We thank Anand Deopurkar and Anthony M. Licata for many useful conversations and ideas.

>>>>

\begin{remark}
\label{rem:numberWigglyPseudotriangulations}
The number~$wp_n$ of wiggly pseudotriangulations of~$n+2$ points is given by
\[
\begin{array}{c|ccccccccc}
n & 1 & 2 & 3 & 4 & 5 & 6 & 7 & 8 & \dots \\
\hline
wp_n & 2 & 14 & 176 & 3232 & 78384 & 2366248 & 85534176 & 3602770400 & \dots
\end{array}
\]
To compute~$wp_n$, denote by $wp_n(x)$ the polynomial where the coefficient of~$x^i$ is the number of wiggly pseudotriangulations of~$n+2$ points with~$i$ \defn{final} internal wiggly arcs (\ie ending at the last point~$n+1$).
For instance,
\begin{align*}
%	wp_0(x) & = 1, \\
	wp_1(x) & = x + 1, \\
	wp_2(x) & = 3 x^3 + 5 x^2 + 4 x + 2, \\
	wp_3(x) & = 15 x^5 + 35 x^4 + 44 x^3 + 40 x^2 + 28 x + 14, \\
	wp_4(x) & = 105 x^7 + 315 x^6 + 520 x^5 + 630 x^4 + 620 x^3 + 514 x^2 + 352 x + 176.
\end{align*}
We invite the reader to check the expressions of~$wp_1(x)$ and~$wp_2(x)$ with \cref{exm:allSmallWigglyPseudotriangulations}.
We use this additional variable~$x$ to obtain a recursive formula for~$wp_n(x)$.
%Indeed, any wiggly pseudotriangulation on~$n+1$ points can be obtained from a wiggly pseudotriangulation on~$n+2$ points by collapsing the points~$n$ and~$n+1$.
Indeed, any wiggly pseudotriangulation of~$n+2$ points can be obtained from a wiggly pseudotriangulation of~$n+1$ points by inserting a single wiggly pseudotriangle with a hinge at~$n$.
%There are two kinds of wiggly pseudotriangulations of~$n+2$ points:
%\item those with the wiggly arc~$(n, n+1, \varnothing, \varnothing)$ are obtained from a wiggly pseudotriangulation of~$n+1$ points by choosing the position of the only corner of~$t_n$ which is not at~$n+1$,
%\item those without the wiggly arc~$(n, n+1, \varnothing, \varnothing)$ are obtained from a wiggly pseudotriangulation of~$n+1$ points by choosing the positions of the two corners of~$t_n$ which are not at~$n+1$.
Each wiggly pseudotriangulation of~$n+1$ points with $i$ final internal wiggly arcs contributes to:
\begin{itemize}
\item $i+2$ wiggly pseudotriangulations of~$n+2$ points with $i+2$ final internal wiggly arcs, containing the wiggly arc~$(n, n+1, \varnothing, \varnothing)$,
\item $j+1$ wiggly pseudotriangulations of~$n+2$ points with $j$ final internal wiggly arcs, not containing the wiggly arc~$(n, n+1, \varnothing, \varnothing)$,  for all~$0 \le j \le i+1$.
\end{itemize}
\pagebreak
\cref{fig:pseudotriangulationsCounting} illustrates this decomposition.
\begin{figure}
\centerline{\includegraphics[scale=1.3]{pseudotriangulationCounting}}
\caption{Some wiggly pseudotriangulations of~$7$ points obtained from the top left wiggly pseudotriangulation of~$6$ points by inserting a wiggly pseudotriangle with hinge at~$5$ as described in \cref{rem:numberWigglyPseudotriangulations}. Those on the first line contain the wiggly arc~$(5, 6, \varnothing, \varnothing)$ and are thus obtained by inserting the head of the wiggly pseudotriangle with hinge at~$5$ in one of the final wiggly arcs. Those on the second line do not contain the wiggly arc~$(5, 6, \varnothing, \varnothing)$ and are thus obtained by inserting the two legs of the wiggly pseudotriangle with hinge at~$5$ in two of the final wiggly arcs. The coloring of the wiggly arcs is supposed to help visualizing in which wiggly arcs the head or legs of the new wiggly pseudotriangle were inserted.}
\label{fig:pseudotriangulationsCounting}
\end{figure}
This translates to the recursive formula %for all~$n \ge 1$,
\begin{align*}
wp_{n}(x) & = \overbracket[.5pt]{x \cdot \frac{\partial}{\partial x} \big( x^2 wp_{n-1}(x) \big)}^{\substack{\text{wiggly pseudotriangulations} \\ \text{containing } (n, n+1, \varnothing, \varnothing)}} + \overbracket[.5pt]{\frac{\partial}{\partial x}\Big( \smash{\frac{wp_{n-1}(1)-x^3wp_{n-1}(x)}{1-x}} \Big)}^{\substack{\qquad\; \text{wiggly pseudotriangulations} \qquad\;  \\ \text{not containing } (n, n+1, \varnothing, \varnothing)}} \\
& = \frac{1}{(1 - x)^2} \big( wp_{n-1}(1) + x^2 (2 x^2 - 2 x - 1) wp_{n-1}(x) + x^4 (x - 1) wp_{n-1}’(x) \big).
\end{align*}
This enables us to quickly compute~$wp_n(x)$ and we obtain the numbers of wiggly pseudotriangulations~$wp_n$ by evaluating~$wp_n(x)$ at~$x = 1$.
%This enables us to quickly compute the generating series of the number of wiggly pseudotriangulations
%\[
%\sum_{n \ge 0} wp_n(1) y^n = 1 + 2 y + 14 y^2 + 176 y^3 + 3232 y^4 + 78384 y^5 + 2366248 y^6 + 85534176 y^7 + 3602770400 y^8 + 
%% 173300710720 y^9 + 9373542317760 y^10 + 563142033172480 y^11 + 37206559614499840 y^12 + 2681213937595142656 y^13 +
%\dots
%\]
\end{remark}

@@

We are now ready to consider the wiggly flip graph and the wiggly complex.
The following definition is illustrated in \cref{fig:wigglyComplex}\,(right) when~$n = 2$.
%
\begin{figure}
\centerline{\includegraphics[scale=1.1]{wigglyComplex}}
\caption{The wiggly complex~$\wigglyComplex_2$ (left) and the wiggly flip graph~$\wigglyFlipGraph_2$ (right). We have embedded the wiggly complex~$\wigglyComplex_2$ in the plane, so that all triangles (including the external one) form a face of~$\wigglyComplex_2$.}
\label{fig:wigglyComplex}
\end{figure}

@@

\begin{remark}
\label{rem:numberWigglyPermutations}
The number~$wp_n$ of wiggly permutations of~$[2n]$ is given by
\[
\begin{array}{c|ccccccccc}
n & 1 & 2 & 3 & 4 & 5 & 6 & 7 & 8 & \dots \\
\hline
wp_n & 2 & 14 & 176 & 3232 & 78384 & 2366248 & 85534176 & 3602770400 & \dots
\end{array}
\]
To compute~$wp_n$, denote by $wp_n(x)$ the polynomial where the coefficient of~$x^i$ is the number of wiggly permutations of~$[2n]$ with $i$ \defn{admissible gaps} (\ie gaps~$\gamma$ between two consecutive positions such that there is no~$j \in [n]$ such that the value~$2j$ appears before the gap~$\gamma$ while the value~$2j-1$ appears after the gap~$\gamma$).
For instance,
\begin{align*}
%	wp_0(x) & = 1, \\
	wp_1(x) & = x + 1, \\
	wp_2(x) & = 3 x^3 + 5 x^2 + 4 x + 2, \\
	wp_3(x) & = 15 x^5 + 35 x^4 + 44 x^3 + 40 x^2 + 28 x + 14, \\
	wp_4(x) & = 105 x^7 + 315 x^6 + 520 x^5 + 630 x^4 + 620 x^3 + 514 x^2 + 352 x + 176.
\end{align*}
We invite the reader to check the expressions of~$wp_1(x)$ and~$wp_2(x)$ with \cref{exm:allSmallWigglyPermutations}.
Any wiggly permutation of~$[2n]$ can be obtained from a wiggly permutation of~$[2n-2]$ by inserting the last values~$2n-1$ and~$2n$.
Each wiggly permutation of~$[2n-2]$ with $i$ admissible gaps contributes to:
\begin{itemize}
\item $i+2$ wiggly permutations of~$[2n]$ with $i+2$ admissible gaps and where~$2n-1$ appears before~$2n$ (they must appear consecutively),
\item $j+1$ wiggly permutations of~$[2n]$ with $j$ admissible gaps and where $2n$ appears before~$2n-1$, for all~$0 \le j \le i+1$.
\end{itemize}
For instance, the wiggly permutation~$78621354$ of~$[8]$ has admissible gaps~$\bullet 7 \bullet 8 \bullet 62135 \bullet 4 \bullet$ and thus gives rise to the wiggly permutations
\[
9\mathrm{X}78621354, \quad 79\mathrm{X}8621354, \quad 789\mathrm{X}621354, \quad 78621359\mathrm{X}4, \quad 786213549\mathrm{X}
\]
of~$[10]$ where the $9$ appears before~$10$, and the wiggly permutations
\begin{align*}
& \mathrm{X}978621354, \quad \mathrm{X}798621354, \quad \mathrm{X}789621354, \quad \mathrm{X}786213594, \quad \mathrm{X}786213549, \\
& 7\mathrm{X}98621354, \quad 7\mathrm{X}89621354, \quad 7\mathrm{X}86213594, \quad 7\mathrm{X}86213549, \quad 78\mathrm{X}9621354, \\
& 78\mathrm{X}6213594, \quad 78\mathrm{X}6213549, \quad 7862135\mathrm{X}94, \quad 7862135\mathrm{X}49, \quad 78621354\mathrm{X}9
\end{align*}
of~$[10]$ where the $10$ appears before~$9$ (for brievety, we have replaced $10$ by $\mathrm{X}$ in the permutations).
We thus obtain the same recursive equation on~$wp_n(x)$ as in \cref{rem:numberWigglyPseudotriangulations}.
\end{remark}

@@

\pagebreak
Finally, consider a cover relation~$\sigma \lessdot \sigma^j$ in the wiggly lattice~$\wigglyLattice_n$ as described in~\cref{lem:wigglyCoverRelation}.
Using the notations of~\cref{lem:wigglyCoverRelation}, observe that
\begin{itemize}
\item $\alpha(\sigma^j, \ell) = \alpha(\sigma, \ell)$ for all~$\ell < i$ or~$\ell > k$,
\item $\alpha(\sigma^j, \ell) = \alpha(\sigma, \ell+k-j)$ for all~$i \le \ell < j$, and
\item $\alpha(\sigma^j, \ell) = \alpha(\sigma, \ell-j+i-1)$ for all~$j < \ell \le k$.
\end{itemize}
Hence, $\Psi(\sigma) \ssm \{\alpha(\sigma, j)\} = \Psi(\sigma^j) \ssm \{\alpha(\sigma^j, j)\}$.
This implies that~$\Psi$ is a directed graph morphism from the Hasse diagram of the wiggly lattice~$\wigglyLattice_n$ to the wiggly increasing flip graph~$\wigglyIncreasingFlipGraph_n$.
As the underlying undirected graphs are both regular of degree~$2n-1$, we conclude that~$\Phi$ and~$\Psi$ induce directed graph isomorphisms between the wiggly increasing flip graph~$\wigglyIncreasingFlipGraph_n$ and the Hasse diagram of the wiggly lattice~$\wigglyLattice_n$.
\end{proof}

@@

\begin{remark}
The map~$\Phi$ sends wiggly pseudotriangulations with $i$ final internal wiggly arcs to wiggly permutations with $i$ admissible gaps, explaining the similarities of \cref{rem:numberWigglyPseudotriangulations,rem:numberWigglyPermutations}.
For instance, the wiggly pseudotriangulations of \cref{fig:pseudotriangulationsCounting} illustrating the description of \cref{rem:numberWigglyPseudotriangulations} correspond to the wiggly permutations
\[
\begin{array}{cccc}
78621354 & 9\mathrm{X}78621354 & 789\mathrm{X}621354 & 78621359\mathrm{X}4 \\
\mathrm{X}786213594 & 7\mathrm{X}89621354 & 78\mathrm{X}6213549 & 7862135\mathrm{X}49
\end{array}
\]
illustrating the description of \cref{rem:numberWigglyPermutations}.
\end{remark}

@@

\pagebreak
\section{Wiggly fan and wigglyhedron}
\label{sec:geometry}

@@

\enlargethispage{.6cm}
Consider now a complete simplicial fan~$\Fan$ of~$\R^d$ with rays~$(\b{r}_v)_{v \in V}$ and cones~$\R_{\ge 0} \b{R}_\triangle$ for~${\triangle \in \Delta}$, where~$\b{R}_\triangle \eqdef \set{\b{r}_v}{v \in \triangle}$ as in \cref{prop:characterizationFan}.
To realize the fan~$\Fan$, one can try to pick a height vector~$\b{h} \eqdef (h_v)_{v \in V} \in \R^V$ and consider the polytope
\(
\polytope{P}_{\b{h}} \eqdef \set{\b{x} \in \R^d}{\dotprod{\b{r}_v}{\b{x}} \le h_v \text{ for all } v \in V}.
\)
The following classical statement characterizes the height vectors~$\b{h}$ for which the fan~$\Fan$ is the normal fan of this polytope~$\polytope{P}_{\b{h}}$.
We borrow the formulation from~\cite[Lem.~2.1]{ChapotonFominZelevinsky}.

@@

\pagebreak
We denote by~$(\b{e}_i)_{i \in [d]}$ the standard basis of~$\R^d$.
For~$I \subseteq [d]$, we define~$\one_I \eqdef \sum_{i \in I} \b{e}_i$, and we often shorten~$\one_{[d]}$ by~$\one_d$.
We denote by~$\HH_d$ the hyperplane of~$\R^d$ defined by the equation~$\dotprod{\b{x}}{\one_d} = 0$, and we denote by~$\pi : \R^d \to \HH_d$ the orthogonal projection on~$\HH_d$, that is~$\pi(\b{x}) \eqdef \b{x} - (\dotprod{\b{x}}{\one_d} / d) \one_d$.

@@

\begin{figure}[h]
	\centering
	\begin{tikzcd}
		1 \arrow[bend left]{r} & 2 \arrow[bend left]{r}\arrow[bend left]{l}& \cdots \arrow[bend left]{r}\arrow[bend left]{l}& m \arrow[bend left]{l}
	\end{tikzcd}
	\caption{The quiver \(Q_m\). The algebra \(A_m\) is a quotient of the path algebra \(kQ_m\).}
	\label{fig:am-quiver}
\end{figure}

@@

This work started during the \emph{Winterbraids XIII: School on low dimensional topology, braid groups and interactions}. We are grateful to the organizers Bérénice Delcroix-Oger, Peter Feller, Vincent Florens, Hoel Queffelec for this fruitful event.
We thank Anand Deopurkar and Anthony M. Licata for many useful conversations and ideas.
Finally, we are grateful to an anonymous referee for many relevant suggestions on the presentation of the paper.
